The journey of my Ph.D. has been an adventurous one, and it is the support of
my advisor, Prof. Purushottam Kulkarni that has made this materialize. I
am grateful to my father, (late) Dr. K. K. Sudevan for inspiring and
encouraging me to take up a challenging Ph.D program at the esteemed
Indian Institute of Technology, Bombay. 
\\
\\
During the journey of my Ph.D., I have made several friends and 
acquaintances who have contributed in ways, known and unknown, 
to my outlook on life. 
I thank my friends and lab mates (Rijurekha Sen, Vijay Gabale,
Uma Sawant, Senthil Nathan, Dhaval Bonde, Viven Rajendra,
Mayank Mishra, Swetha Srinivasan, Prashima Sharma,
Prajakta Patil, Mukulika Maity, to name a few),
for the innumerable discussions, both technical and personal.
I thank my friend and roommate, Deepti Shenai, for her continued
encouragement and support. 
\\
\\
I also thank fellow members on the hostel student council,
during my term as General Secretary, especially Bandana Singha,
Rohini Karandikar, Rijurekha Sen, Urbashi Sarkar and Neelam Rathore.
I also acknowledge my friends and
colleagues, Abhisekh Sankaran and Karthik Ramachandra, for
engaging with my musical side.
%I am indebted especially to Senthil Nathan, who has offered me
%consistently good technical feedback and encouragement.
\\
\\
Last but not the least, I am grateful to my mother, Mrs. Usha Sudevan, 
my husband, Amey Gavand, and my
parents-in-law, Dr. Narayan Balaram Gavand and Mrs. Neelam Gavand,
for their undying support and encouragement. Without their support,
we would not be here today. Thank you.
