\paragraph{Conference Proceedings}

\begin{enumerate}
    \item \textit{DRIVE: Using Implicit Caching Hints to achieve Disk I/O Reduction in Virtualized Environments}. Proceedings of the 21st International Conference on High Performance Computing (HiPC), 2014. Sujesha Sudevalayam, Purushottam Kulkarni, Rahul Balani and Akshat Verma. 
    \item \textit{Affinity-aware Modeling of CPU Usage for Provisioning Virtualized Applications}. Proceedings of the 4th International Conference on Cloud Computing (CLOUD), 2011. Sujesha Sudevalayam and Purushottam Kulkarni.
\end{enumerate}

\paragraph{Journal Publications}

\begin{enumerate}
    \item \textit{Affinity-aware Modeling of CPU Usage with Communicating Virtual Machines}. Journal of Systems and Software (JSS), 2013. Sujesha Sudevalayam, Purushottam Kulkarni.
    \item \textit{Energy Harvesting Sensor Nodes: Survey \& Implications}. IEEE Communications Surveys and Tutorials 2011. Sujesha Sudevalayam and Purushottam Kulkarni.
\end{enumerate}

\paragraph{Technical Reports}

\begin{enumerate}
\item \textit{CONFIDE: Content-based Fixed-sized I/O Deduplication}. Technical Report, IIT Bombay. Sujesha Sudevalayam, Purushottam Kulkarni, Rahul Balani and Akshat Verma. IITB/CSE/2014/April/60, TR-CSE-2014-60.
\item \textit{Affinity-aware Modeling of CPU Usage for Provisioning Virtualized Applications}. Technical Report, IIT Bombay. Sujesha Sudevalayam and Purushottam Kulkarni. IITB/CSE/2011/February/34, TR-CSE-2011-34.
\item \textit{Colocation-aware Modeling of CPU Usage for P2V Transitioning Applications}. Technical Report, IIT Bombay. Sujesha Sudevalayam. IITB/CSE/2014/March/59, TR-CSE-2014-59.
\item \textit{Balancing Response Time and CPU allocation in Virtualized Data Centers using Optimal Controllers}. Technical Report, IIT Bombay. Varsha Apte, Purushottam Kulkarni, Sujesha Sudevalayam and Piyush Masrani. IITB/CSE/2010/April/29, TR-CSE-2010-29.
\item \textit{Energy Harvesting Sensor Nodes: Survey and Implications}. Technical Report, IIT Bombay. Sujesha Sudevalayam and Purushottam Kulkarni. IITB/CSE/2008/December/19, TR-CSE-2008-19.
\end{enumerate}
